### 2013-03-10 doc/sosp13/memory.tex
sosp repo. put in sosp submission guidelines. new title
lots of edits. more coming

Applications written for Grappa utilize two forms of memory: local and
global.  Local memory is local to a single core in the system.  Accesses occur
through conventional pointers.  The compiler emits an access and the
memory is manipulated directly.  Applications use local accesses for a
number of things in Grappa: the stack associated with a task, accesses
to localized global memory in caches (see below), and accesses to
debugging infrastructure that is local to each system node.  Local
pointers cannot access memory on other cores, and are valid only on
their home core.

@@

\begin{figure}[htbp]
  \begin{center}
    \begin{description}\small
      \item[ \texttt{ global\_address global\_malloc( size )} ] \hfill \\
      \item[ \texttt{ global\_free( global\_address )} ] \hfill \\
        Allocates and frees memory in the global heap
      \item[ \texttt{ delegate\_read( global\_address, local\_var )} ] 
      \item[ \texttt{ delegate\_write( global\_address, local\_var )} ] %\vspace{-2ex}
      \item[ \texttt{ delegate\_cas( global\_address, local\_var )} ] %\vspace{-2ex}
      \item[ \texttt{ delegate\_fetch\_inc( global\_address, local\_var )} ] %\vspace{-2ex} 
\hfill \\
        Performs a memory operation at the home core of a global address
      \item[ \texttt{ cache\_acquire( global\_address, local\_buf, \{RO,RW,WO\})} ]
      \item[ \texttt{ cache\_release( global\_address, local\_buf )} ] %\vspace{-2ex} 
\hfill \\
        Perform cache operations \TODO{expand}
    \end{description}
    \begin{minipage}{0.95\columnwidth}
      \caption{\label{fig:accessing-memory} Grappa API: accessing memory} %{-4ex}}
    \end{minipage}
    %\vspace{-3ex}
  \end{center}
\end{figure}

>>>>

Applications written for Grappa utilize two forms of memory: local and global.
Local memory is local to a single core \comment{core or processor/node?} in
the system. Accesses occur through conventional pointers. The compiler emits
an access and the memory is manipulated directly. Applications use local
accesses for a number of things in Grappa: the stack associated with a task,
accesses to localized global memory in caches (see below), and accesses to
debugging infrastructure that is local to each system node. Local pointers
cannot access memory on other cores, and are valid only on their home core.

@@

\begin{figure}[htbp]
  \begin{center}
    \begin{minipage}{\columnwidth}
	\small
	\begin{tabular}{l}
      	\texttt{\scriptsize global\_address global\_malloc( size )} \\
      	\texttt{\scriptsize global\_free( global\_address )} \\ \hline 
      	Allocation in the global heap \\ \\
      	\texttt{\scriptsize delegate\_read( global\_address, local\_var )}  \\
      	\texttt{\scriptsize delegate\_write( global\_address, local\_var )} \\
      	\texttt{\scriptsize delegate\_cas( global\_address, local\_var )} \\
      	\texttt{\scriptsize delegate\_fetch\_inc( global\_address, local\_var )} \\ \hline
      	Memory operations at the home core of a global address \\ \\
      	\texttt{\scriptsize cache\_acquire( global\_address, local\_buf, \{RO,RW,WO\})} \\
      	\texttt{\scriptsize cache\_release( global\_address, local\_buf )} \\ \hline
		Perform cache operations to acquire/release global data.  \\
		Acquire copies all data to local node and returns a pointer. \\ 	
		Release optionally copies data back to global memory. \\
	\end{tabular}
      \caption{\label{fig:accessing-memory} Grappa API: accessing memory}     \end{minipage}
  \end{center}
\end{figure}

\comment{
    \begin{description}\small
      \item[ \texttt{ global\_address global\_malloc( size )} ] \hfill \\
      \item[ \texttt{ global\_free( global\_address )} ] \hfill \\
        Allocates and frees memory in the global heap
      \item[ \texttt{ delegate\_read( global\_address, local\_var )} ] 
      \item[ \texttt{ delegate\_write( global\_address, local\_var )} ] 
      \item[ \texttt{ delegate\_cas( global\_address, local\_var )} ] 
      \item[ \texttt{ delegate\_fetch\_inc( global\_address, local\_var )} ] 
\hfill \\
        Performs a memory operation at the home core of a global address
      \item[ \texttt{ cache\_acquire( global\_address, local\_buf, \{RO,RW,WO\})} ]
      \item[ \texttt{ cache\_release( global\_address, local\_buf )} ] 
\hfill \\
Perform cache operations to acquire/release global data.  Acquire, returns a local pointer after all data has been copied to the local node.  Release, optionally writes data back to global memory and frees up management resources.
    \end{description}
}

### 2013-03-10 doc/sosp13/overview.tex
edits all over. some eval suggestions
lots of edits. more coming

\item [Tasking system.] Our tasking system supports lightweight
multithreading to tolerate communication latency and global distributed
workstealing (i.e., tasks can be stolen from any node in the system), which
provides automated load balancing.

>>>>

\item [Tasking system.] Our tasking system supports lightweight multithreading
to tolerate communication latency and global distributed work-stealing (i.e.,
tasks can be stolen from any node in the system), which provides automated
load balancing. As discussed earlier, the reason we built our own custom
tasking system is because of the need for large number of active tasks and the
resulting need for careful task context prefetching.

% \TODO{say why we built our own: need to support large number of contexts and just the contexts themselves wont fit in the cache, so being able to prefetch task contexts is very important.}

### 2013-03-10 doc/sosp13/tasks.tex
lots of edits. more coming

During execution, a task yields control of its core whenever it performs
a long-latency operation, allowing the processor to remain busy while
waiting for the operation to complete.  In addition,  a programmer can
direct scheduling explicitly via the Grappa API calls shown
in Figure~\ref{fig:scheduling}.
To minimize yield overhead, the Grappa scheduler operates entirely in
user-space and does little more than store register state of one task and load
that of another. Context switch times are as low as 38ns even when switching
amongst thousands of tasks.

@@

\subsection{Expressing parallelism}

@@

Third, the programmer can invoke a parallel for loop, provided that
the trip count is known at loop entry. The programmer
specifies a function pointer along with start and end indices and an
optional threshold to control parallel overhead. Grappa does {\em
recursive decomposition} of iterations, similar to Cilk's cilk\_for
construct~\cite {cilkforimplementation}.\comment{could only find slides
page6 of
http://www.clear.rice.edu/comp422/lecture-notes/comp422-2012-Lecture5-
Cilk++.pdf    ---- Good enough.  Cite it.  -Mark} and TBB's {\tt parallel\_for}~\cite{intel_tbb}.  It generates a
logarithmically-deep tree of tasks, stopping to execute the loop body
when the number of iterations is below the required threshold.

@@

\subsection{Implementation tasks}

\paragraph{Tasks and workers} Grappa tasks are 32-byte entities: a
64-bit function pointer plus three 64-bit arguments. We use three
arguments because tasks are often generated via a parallel loop
decomposition, in which each task needs three kinds of data.
\begin{description}
\item[Function pointer] addresses the routine to run. We configure the system nodes to disable address randomization, so that function pointers are valid across process images.
\item[Private argument] often a loop index.
\item[Shared argument] typically data shared amongst a group of tasks, or the number of loop iterations to be performed by this task
\item[Synchronization argument] often used to determine
  when all tasks that are part of a loop have finished; a global pointer to a synchronization object allocated at the core that spawned a group of tasks.
\end{description}
While these are the most common uses of the three task arguments, they
are treated as arbitrary 64-bit values in the runtime, and can be used
for any purpose.

@@

Each scheduler has three queues:
\begin{description}
\item{\bf Ready worker queue} This is a FIFO queue of tasks that are
  matched with workers and are ready to execute.
\item{\bf Private task queue} This is a FIFO queue of tasks that must run on this core.
\item{\bf Public task queue} This is a LIFO queue of tasks that are
  waiting to be matched with workers. It is a local partition of a shared
  task pool.
\end{description}

>>>>

During execution, a task yields control of the processor core whenever it
performs a long-latency operation, allowing the processor to remain busy while
waiting for the operation to complete. In addition, a programmer can direct
scheduling explicitly via the Grappa API calls shown in
Figure~\ref{fig:scheduling}. To minimize yield overhead, the Grappa scheduler
operates entirely in user-space and does little more than store register state
of one task and load that of another. Context switch times are as low as 38ns
even when switching amongst thousands of tasks.

@@

\subsection{Expressing Parallelism}

@@

Third, the programmer can invoke a parallel for loop, provided that the trip
count is known at loop entry. The programmer specifies a function pointer
along with start and end indices and an optional threshold to control parallel
overhead. Grappa does {\em recursive decomposition} of iterations, similar to
Cilk's cilk\_for construct~\cite {cilkforimplementation}, and TBB's {\tt
parallel\_for}~\cite{intel_tbb}. It generates a logarithmically-deep tree of
tasks, stopping to execute the loop body when the number of iterations is
below the required threshold.

@@

\subsection{Task Support Implementation}

\paragraph{Tasks and workers} Grappa tasks are 32-byte entities: a 64-bit
function pointer plus three 64-bit arguments. We use three arguments because
tasks are often generated via a parallel loop decomposition, in which each
task needs three kinds of data. Thus, tasks include:

\noindent {\it function pointer},
or the addresses of the routine to run --- we configure the system nodes to
disable address randomization, so that function pointers are valid across
process images;

\noindent {\it private argument}, often a loop index;

\noindent {\it shared
argument}, typically data shared amongst a group of tasks, or the number of
loop iterations to be performed by this task;

\noindent {\it synchronization argument}, often used to determine when all
tasks that are part of a loop have finished, in the form of q global pointer
to a synchronization object allocated at the core that spawned a group of
tasks.

While these are the most common uses of the three task arguments, they are
treated as arbitrary 64-bit values in the runtime, and can be used for any
purpose.

@@

Each scheduler has three queues:

\noindent {\it ready worker queue}, a FIFO queue of tasks that are
  matched with workers and are ready to execute;

\noindent {\it private task queue}, a FIFO queue of tasks that must run on this core;

\noindent {\it public task queue},  a LIFO queue of tasks that are
  waiting to be matched with workers. It is a local partition of a shared
  task pool.
